---
title: "CH3 Analysis"
author: "R Pienaar"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Libraries

```{r}

library(ggplot2)
library(purrr)
library(lme4)
library(performance)
library(glmmTMB)
library(tidyverse)
library(ggpubr)
library(MuMIn)
options(scipen=999)

common_theme <- theme(
  text = element_text(size = 16, color = "black", family = "serif"),
  axis.title = element_text(size = 16, color = "black", family = "serif"),
  axis.text = element_text(size = 16, color = "black", family = "serif"),
  legend.text = element_text(size = 16, color = "black", family = "serif"),
  legend.title = element_text(size = 16, color = "black", family = "serif"),
  plot.title = element_text(size = 16, color = "black", family = "serif", hjust = 0.5)
)

```


Lets explore the data starting with the Great Basin

Load file

```{r}

gbdata <- read.csv("GreatBasin_Data.csv")

ggplot(data = gbdata, aes(x= Year, y= rich))+
  stat_summary(geom = "line", fun = "mean")

ggplot(data = gbdata, aes(x= Year, y = simp_div))+
  stat_summary(geom = "line", fun = "mean")

ggplot(data = gbdata, aes(x= Year, y = shan_div))+
  stat_summary(geom = "line", fun = "mean")

```

Our first research question was on the effects of fire and rodent exclusion so I am going to start with that

```{r}

head(gbdata)

# Create Fire column with the first letter of Plot
gbdata$Fire <- substr(gbdata$Plot, 1, 1)

# Create Rodents column with the second letter of Plot
gbdata$Rodents <- substr(gbdata$Plot, 2, 2)

summary(gbdata$Rodents)

gbdata$Block <- as.factor(gbdata$Block)
gbdata$Plot <- as.factor(gbdata$Plot)
gbdata$Fire <- as.factor(gbdata$Fire)
gbdata$Fire <- factor(gbdata$Fire, levels = c("U", "B"))
gbdata$Rodents <- as.factor(gbdata$Rodents)
gbdata$Rodents <- factor(gbdata$Rodents, levels = c("S", "N"))

# Recode Fire column
gbdata$Fire <- recode(gbdata$Fire, 
                      "B" = "Burned", 
                      "U" = "Unburned")

# Recode Rodents column
gbdata$Rodents <- recode(gbdata$Rodents, 
                         "N" = "Rodents Excluded", 
                         "S" = "Rodents Present")


```

Start with Richness

```{r}

acf(gbdata$rich)

# I have several resposne variables I could use, i will start with richness, diversity, BRTE

# Richness anova
rich.lm <- glmmTMB(data = gbdata, rich ~ Fire * Rodents +(1|Block) + ar1(Block + 0 | Year))
summary(rich.lm)
r2(rich.lm)
# Results 
# Burning increases plant species richness in the Great Basin (B = 0.39, SE = 0.15, p = 0.009)
# Rodent exclusion decreases plant species richness in the Great Basin (B = -0.26, SE = 0.15, p 0.075)
# There is a significant interactive effect between burning and rodent exclusion (B = 0.48, SE = 0.21, p = 0.022)

# Now plot it

rich.plot <- ggplot(data = gbdata, aes(x = Year, y = rich, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(0,7)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin Species Richness") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))

rich.plot

```

Lets do Simpsons diversity next

```{r}

acf(gbdata$simp_div)
# 
simp.lm <- glmmTMB(data = gbdata, simp_div ~ Fire * Rodents + (1|Block))
summary(simp.lm)
# Results 
# No significant effects of fire or rodents on simpsons diversity

# Now plot it
hist(gbdata$simp_div)

simp.plot <- ggplot(data = gbdata, aes(x = Year, y = simp_div, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin Simpson's Diversity") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))


simp.plot

```

Now Shannon's Diversity

```{r}

acf(gbdata$shan_div)
# 
shan.lm <- glmmTMB(data = gbdata, shan_div ~ Fire * Rodents +(1|Block))
summary(shan.lm)
# Results 
# Burning slightly increased Shannons diversity in the Great Basin (B = 0.06, SE = 0.03, p = 0.09)

# Now plot it

shan.plot <- ggplot(data = gbdata, aes(x = Year, y = shan_div, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  coord_cartesian(ylim = c(0,1.5)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin Shannon's Diversity") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))

shan.plot

```

Now lets look at cheatgrass density

```{r}

# 
brte.lm <- glmmTMB(data = gbdata, BRTE ~ Fire * Rodents +(1|Block))
summary(brte.lm)

acf(gbdata$BRTE) ## No autocorrellation


# Now plot it

brte.plot <- ggplot(data = gbdata, aes(x = Year, y = BRTE, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin \n Bromus tectorum \n Density (m^2)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))

brte.plot
summary(gbdata)

```

Squirrel tail

```{r}

# 
elel.lm <- glmmTMB(data = gbdata, ELEL ~ Fire * Rodents +(1|Block))
summary(elel.lm)
# Results 
# the interaction between fire and rodents was significant (B = 4.02, SE = 2.32, p = 0.08)

# Now plot it

elel.plot <- ggplot(data = gbdata, aes(x = Year, y = ELEL, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin \n Elymus elymoides \n Density (m^2)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))


elel.plot

```

Cete

```{r}


acf(gbdata$CETE)
# 
cete.lm <- glmmTMB(data = gbdata, CETE ~ Fire * Rodents +(1|Block))
summary(cete.lm)
# Results 
# Rodents exclusion decreased CETE density (B = -69.3, SE = 34.69, p = 0.046)
# the interaction between fire and rodents was significant (B = -86.05, SE = 50.26, p = 0.09)

# Now plot it

cete.plot <- ggplot(data = gbdata, aes(x = Year, y = CETE, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin \n Ceratocephala testiculata \n Density (m^2)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))


cete.plot

```

ALAL

```{r}

acf(gbdata$ALAL)

alal.lm <- glmmTMB(data = gbdata, ALAL ~ Fire * Rodents +(1|Block))
summary(alal.lm)
# Results 
# the interaction between fire and rodents was significant (B = 135.93, SE = 15.15, p < 0.0001)

# Now plot it

alal.plot <- ggplot(data = gbdata, aes(x = Year, y = ALAL, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin \n Allysium allysoides \n Density(m^2)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))


alal.plot

```

Lets plot climate
Total Annual Precipitation

```{r}

ggplot(data = gbdata, aes(x = Year, y = TAP)) +
  geom_line()+
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin \n Total Annual Preciptition (mm)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) 

```

Mean temperature

```{r}

ggplot(data = gbdata, aes(x = Year, y = MAT)) +
  geom_line()+
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Great Basin \n Mean Annual Temperature (C)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) 

```

Now lets try that with the mojave data

```{r}

mojdata <- read.csv("Mojave_Data.csv")
head(mojdata)

summary(mojdata)
```

```{r}

# Fix that one stupid plot
mojdata$Plot <- recode(mojdata$Plot, "UN " = "UN")
mojdata$Plot <- as.character(mojdata$Plot)
mojdata$Plot <- as.factor(mojdata$Plot)

# Create Fire column with the first letter of Plot
mojdata$Fire <- substr(mojdata$Plot, 1, 1)

# Create Rodents column with the second letter of Plot
mojdata$Rodents <- substr(mojdata$Plot, 2, 2)

mojdata$Block <- as.factor(mojdata$Block)
mojdata$Plot <- as.factor(mojdata$Plot)
mojdata$Fire <- as.factor(mojdata$Fire)
mojdata$Fire <- factor(mojdata$Fire, levels = c("U", "B"))
mojdata$Rodents <- as.factor(mojdata$Rodents)
mojdata$Rodents <- factor(mojdata$Rodents, levels = c("S", "N"))

# Recode Fire column
mojdata$Fire <- recode(mojdata$Fire, 
                      "B" = "Burned", 
                      "U" = "Unburned")

# Recode Rodents column
mojdata$Rodents <- recode(mojdata$Rodents, 
                         "N" = "Rodents Excluded", 
                         "S" = "Rodents Present")


```

Start with Richness

```{r}

acf(mojdata$rich)

# I have several resposne variables I could use, i will start with richness, diversity, BRTE

# Richness anova
moj.rich.lm <- glmmTMB(data = mojdata, rich ~ Fire * Rodents +(1|Block) + ar1(Block + 0 | Year))
summary(moj.rich.lm)
# Results 
# Burning increases plant species richness in the Mojave (B = 0.40, SE = 0.15, p = 0.06)


# Now plot it

moj.rich.plot <- ggplot(data = mojdata, aes(x = Year, y = rich, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9), linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Species Richness") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=16, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=16, family = "serif"),
        axis.title.x  = element_text(size=16, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=16, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=16, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=16, family = "serif", colour = "black"),
        legend.text = element_text(size=16, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                    labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access"))


moj.rich.plot

# Percent change for paper
aggregate(rich ~ Fire, data = mojdata, mean)
# Unburned	6.488889	
# Burned	6.901944
 (6.901944 - 6.488889)/6.488889*100

#Richness increased by 6.365574

# Percent change by year for paper
aggregate(rich ~ Fire + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Fire, values_from = rich) %>%
  mutate(perc_diff = abs((Burned - Unburned) / Unburned) * 100)

# Percent change for paper
aggregate(rich ~ Rodents, data = mojdata, mean)
# Present 6.706944	
# Absent 6.683889

#Richness increased by 6.365574

# Percent change by year for paper
aggregate(rich ~ Rodents + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Rodents, values_from = rich) %>%
  mutate(perc_diff = abs((`Rodents Present` - `Rodents Excluded`) / `Rodents Excluded`) * 100)


```

```{r}

acf(mojdata$simp_div)

# 
moj.simp.lm <- glmmTMB(data = mojdata, simp_div ~ Fire * Rodents+ Year + ar1(Block + 0 | Year))

summary(moj.simp.lm)
# Results 
# Rodent exclusion increased simpson's diversity in the Mojave (B = 0.06, SE = 0.02, p = 0.0006)

# Now plot it

moj.simp.plot <- ggplot(data = mojdata, aes(x = Year, y = simp_div, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Simpson's Diversity") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))


moj.simp.plot

# Percent change for paper
aggregate(rich ~ Fire, data = mojdata, mean)
# Unburned	6.488889	
# Burned	6.901944
 (6.901944 - 6.488889)/6.488889*100

#Richness increased by 6.365574

# Percent change by year for paper
aggregate(rich ~ Fire + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Fire, values_from = rich) %>%
  mutate(perc_diff = abs((Burned - Unburned) / Unburned) * 100)

# Percent change for paper
aggregate(rich ~ Rodents, data = mojdata, mean)
# Present 6.706944	
# Absent 6.683889

#Richness increased by 6.365574

# Percent change by year for paper
aggregate(rich ~ Rodents + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Rodents, values_from = rich) %>%
  mutate(perc_diff = abs((`Rodents Present` - `Rodents Excluded`) / `Rodents Excluded`) * 100)

```

```{r}

acf(mojdata$shan_div)
# 
moj.shan.lm <- glmmTMB(data = mojdata, shan_div ~ Fire * Rodents+ Year + ar1(Block + 0 | Year))
summary(moj.shan.lm)
VarCorr(moj.shan.lm)
# Results 
# Burning slightly increased Shannons diversity in the Mojave (B = 0.08, SE = 0.04, p = 0.043)
# Rodent exclusion increased Shannons diversity in the Mojave (B = 0.13, SE = 0.04, p = 0.0003)

# Now plot it

moj.shan.plot <- ggplot(data = mojdata, aes(x = Year, y = shan_div, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9), linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Shannon's Diversity") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=16, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=16, family = "serif"),
        axis.title.x  = element_text(size=16, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=16, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=16, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=16, family = "serif", colour = "black"),
        legend.text = element_text(size=16, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                    labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access"))


moj.shan.plot

# Percent change for paper
aggregate(shan_div ~ Fire, data = mojdata, mean)
# Unburned	0.7184890		
# Burned	0.7953773	
 (0.7953773	 - 0.7184890	)/0.7184890	*100

#Shannon's Diversity increased by 10.70139 %

# Percent change by year for paper
aggregate(shan_div ~ Fire + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Fire, values_from = shan_div) %>%
  mutate(perc_diff = abs((Burned - Unburned) / Unburned) * 100)

# Percent change for paper
aggregate(shan_div ~ Rodents, data = mojdata, mean)
# Present 0.7115092	
# Absent 0.8023571
(0.8023571- 0.7115092	)/0.7115092	*100

# Rodent exclusion increased diversity by 12.76834 %

#Richness increased by 6.365574

# Percent change by year for paper
aggregate(shan_div ~ Rodents + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Rodents, values_from = shan_div) %>%
  mutate(perc_diff = abs((`Rodents Present` - `Rodents Excluded`) / `Rodents Excluded`) * 100)

```


Total Density


```{r}

mojdata$totaldensity <- rowSums(mojdata[5:63])

acf(mojdata$totaldensity)

# BRRU instead of BRTE for the Mojave
moj.td.lm <- glmmTMB(data = mojdata, totaldensity ~ Fire * Rodents + ar1(Block + 0 | Year)) 
summary(moj.td.lm)
# Results 
# No significant effects

# Now plot it

moj.td.plot <- ggplot(data = mojdata, aes(x = Year, y = totaldensity, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9), linewidth =1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab(bquote("Total Plant Density  " (m^2))) +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=16, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=16, family = "serif"),
        axis.title.x  = element_text(size=16, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=16, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=16, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=16, family = "serif", colour = "black"),
        legend.text = element_text(size=16, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                    labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access"))

moj.td.plot

# Percent change for paper
aggregate(totaldensity ~ Fire, data = mojdata, mean)
# Unburned	864.5775		
# Burned	1228.5234	
 (1228.5234		 - 864.5775	)/864.5775	*100

#Total density increased by 42.09523 %

# Percent change by year for paper
aggregate(totaldensity ~ Fire + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Fire, values_from = totaldensity) %>%
  mutate(perc_diff = abs((Burned - Unburned) / Unburned) * 100)

# Percent change for paper
aggregate(totaldensity ~ Rodents, data = mojdata, mean)
# Present 1088.032	
# Absent 1005.068	
(1005.068	- 1088.032		)/1088.032		*100

# Rodent exclusion decreased  density by 7.625143%

#Richness increased by 6.365574

# Percent change by year for paper
aggregate(totaldensity ~ Rodents + Year, data = mojdata, mean) |>
  pivot_wider(names_from = Rodents, values_from = totaldensity) %>%
  mutate(perc_diff = abs((`Rodents Present` - `Rodents Excluded`) / `Rodents Excluded`) * 100)

```


BRRU

```{r}

acf(mojdata$BRRU)

# BRRU instead of BRTE for the Mojave
moj.brru.lm <- glmmTMB(data = mojdata, BRRU ~ Fire * Rodents +(1|Block)+ ar1(Block + 0 | Year)) 
summary(moj.brru.lm)
# Results 
# No significant effects

# Now plot it

moj.brru.plot <- ggplot(data = mojdata, aes(x = Year, y = BRRU, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab(bquote("Bromus rubens Density  " (m^2))) +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))

moj.brru.plot

```

ERCI

```{r}


# BRRU instead of BRTE for the Mojave
moj.erci.lm <- glmmTMB(data = mojdata, ERCI ~ Fire * Rodents +(1|Block)) 
summary(moj.erci.lm)
# Results 
# Fire increased ERCI in the mojave (B = 34.64, SE = 4.42, p < 0.0001)
# There is also a significant interactin (B = 10.31, SE = 6.22, = 0.0979)

# Now plot it

moj.erci.plot <- ggplot(data = mojdata, aes(x = Year, y = ERCI, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Mojave \n Erodium cicutarium \n Density (m^2)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))
moj.erci.plot


```

SCAR
```{r}

acf(mojdata$SCAR)

# BRRU instead of BRTE for the Mojave
moj.scar.lm <- glmmTMB(data = mojdata, SCAR ~ Fire * Rodents +(1|Block)+ ar1(Block + 0 | Year)) 
summary(moj.scar.lm)
# Results 
# Fire increased SCAR in the mojave (B = 60.43, SE = 15.39, p < 0.0001)
# Rodent exclusion decreased SCAR (B = -51016, SE = 15.02, p = 0.0007)

# Now plot it

moj.scar.plot <- ggplot(data = mojdata, aes(x = Year, y = SCAR, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab(bquote("Schismus arabicus Density  " (m^2))) +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))

moj.scar.plot


```

ASNU

```{r}


# BRRU instead of BRTE for the Mojave
moj.asnu.lm <- glmmTMB(data = mojdata, ASNU ~ Fire * Rodents +(1|Block)) 
summary(moj.asnu.lm)
# Results 
# Fire increased ASNU in the mojave (B = 11.57, SE = 2.98, p = 0.0001)
# Rodent exclusion increased ASNU (B = 12.27, SE = 2.9, p < 0.0001)

# Now plot it

moj.asnu.plot <- ggplot(data = mojdata, aes(x = Year, y = ASNU, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9), linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Mojave \n Astragalus nuttallianus \n Density (m^2)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=16, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=16, family = "serif"),
        axis.title.x  = element_text(size=16, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=16, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=16, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=16, family = "serif", colour = "black"),
        legend.text = element_text(size=16, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))

moj.asnu.plot


```

Mojave climate plots

```{r}

ggplot(data = mojdata, aes(x = Year, y = TAP)) +
  geom_line()+
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Mojave \n Total Annual Preciptition (mm)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) 

```

```{r}

ggplot(data = mojdata, aes(x = Year, y = MAT)) +
  geom_line()+
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Mojave \n Mean Annual Temperature (C)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) 

```

Rodent abudnance

```{r}

ggplot(data = mojdata, aes(x = Year, y = Rodent.Abundance, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9)) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Mojave \n rodeny \n Density (m^2)") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text = element_text(size=12, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24))

```
Do the cover plot from the mcc dataframe

```{r}

mcc <- read.csv("Moj_Cover.csv")
acf(mcc$Total.Cover)
mcc$Fire <- factor(mcc$Fire, levels = c("Unburned", "Burned"))
mcc$Rodents <- factor(mcc$Rodents, levels = c("Rodents Present", "Rodents Excluded"))
mcc$Block <- as.factor(mcc$Block)

#Also bring in GB
gbcc <- read.csv("rushcoverpercent.csv")
summary(gbcc)
gbcc$Block <- as.factor(gbcc$Block)
gbcc$Plot <- as.factor(gbcc$Plot)
gbcc <- gbcc |>
  rename(Year = DATE)
#merge gbdata and gbcc
gbdenscov <- merge(gbdata, gbcc, by = c("Year", "Block", "Plot"))
write.csv(gbdenscov, "gbdenscov.csv")
# BRRU instead of BRTE for the Mojave
moj.cover.lm <- glmmTMB(data = mcc, Total.Cover ~ Fire * Rodents+ Year +ar1(Block +0 |Year)) 
VarCorr(moj.cover.lm)
summary(moj.cover.lm)

mcc %>%
  filter(Year == 2019) %>%
  summarise(avg_cover = mean(Total.Cover, na.rm = TRUE))
# Now plot it

moj.cover.plot <- ggplot(data = mcc, aes(x = Year, y = Total.Cover, fill = Plot, shape = Plot, colour = Plot)) +
  stat_summary(geom = "point", fun = 'mean', position = position_dodge(width = 0.9), size = 3) +
  stat_summary(geom = "line", fun = "mean", position = position_dodge(width = 0.9), linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               position = position_dodge(width = 0.9), 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Total Plant Cover %") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=16, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=16, family = "serif"),
        axis.title.x  = element_text(size=16, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=16, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=16, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=16, family = "serif", colour = "black"),
        legend.text = element_text(size=16, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
  scale_fill_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                    labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_color_manual(values = c("BN" = "red", "BS" = "black", "UN" = "blue", "US" = "green3"),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access")) +
  scale_shape_manual(values = c("BN" = 21, "UN"= 21, "BS" = 24, "US" = 24),
                     labels = c("BN" = "Burned Rodent Exclusion", "BS" = "Burned Rodent Access", "UN" = "Unburned Rodent Exclusion", "US" = "Unburned Rodent Access"))

moj.cover.plot

# Percent change for paper
aggregate(Total.Cover ~ Fire, data = mcc, mean)
# Unburned	60.84322		
# Burned	67.03767
 (67.03767	 - 60.84322	)/60.84322		*100

#Richness increased by 10.181%

# Percent change by year for paper
aggregate(Total.Cover ~ Fire + Year, data = mcc, mean) |>
  pivot_wider(names_from = Fire, values_from = Total.Cover) %>%
  mutate(perc_diff = abs((Burned - Unburned) / Unburned) * 100)


# Percent change for paper
aggregate(Total.Cover ~ Rodents, data = mcc, mean)
# Presemt	64.49100		
# Burned	63.38989	
 (63.38989 - 64.49100)/64.49100		*100

# -1.707386 %

# Percent change by year for paper
aggregate(Total.Cover ~ Rodents + Year, data = mcc, mean) |>
  pivot_wider(names_from = Rodents, values_from = Total.Cover) %>%
  mutate(perc_diff = abs((`Rodents Present` - `Rodents Excluded`) / `Rodents Excluded`) * 100)

```


Put all those plots together

```{r}

ggarrange(rich.plot, moj.rich.plot,
          simp.plot, moj.simp.plot,
          shan.plot, moj.shan.plot,
          ncol = 2, nrow = 3,
          common.legend = TRUE,
          legend = "right")

ggarrange(moj.td.plot, moj.rich.plot,
          moj.cover.plot, moj.shan.plot,
          ncol = 1, nrow = 4,
          common.legend = TRUE,
          legend = "right",
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv")

ggarrange(moj.td.plot, moj.rich.plot,
          moj.cover.plot, moj.shan.plot,
          ncol = 2, nrow = 2,
          common.legend = TRUE,
          legend = "right",
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv")

ggarrange(
          moj.shan.plot, moj.cover.plot,
          ncol = 2, nrow = 1,
          common.legend = TRUE,
          legend = "right",
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv")

```

put all the species plots together

```{r}

ggarrange(brte.plot, moj.brru.plot,
          elel.plot, moj.scar.plot,
          alal.plot, moj.erci.plot,
          cete.plot, moj.asnu.plot,
          ncol = 2, nrow = 4,
          common.legend = TRUE,
          legend = "right")

ggarrange(moj.brru.plot,
          moj.scar.plot,
           ncol = 1, nrow = 2,
          common.legend = TRUE,
          legend = "right")


```

just invasive species plots

```{r}

ggarrange(brte.plot, moj.brru.plot,
          cete.plot, moj.scar.plot,
          alal.plot, 
          ncol = 2, nrow = 3,
          common.legend = TRUE,
          legend = "right")




```

Bar graphs for plot effects

```{r}


g1 <- ggplot(data = mcc, aes(x = Fire, y = Total.Cover, fill = Rodents, colour = Rodents)) +
  stat_summary(geom = "bar", fun = mean, position = position_dodge(width = 0.9)) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.9), width = 0.2, color = "black") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,100)) +
  common_theme +  
  ylab("Total Plant Cover %")+
  scale_fill_manual(values = c("Rodents Present" = "white", "Rodents Excluded" = "gray35"))+
  scale_color_manual(values = c("Rodents Present" = "black", "Rodents Excluded" = "black"))  # Black outlines

g1

g2 <- ggplot(data = mojdata, aes(x = Fire, y = shan_div, fill = Rodents, color = Rodents)) +
  stat_summary(geom = "bar", fun = mean, position = position_dodge(width = 0.9)) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 0.9), width = 0.2, color = "black") +
  theme_classic() +
  scale_y_continuous(expand = c(0,0), limits = c(0,2)) +
  coord_cartesian(ylim = c(0,1))+
  common_theme +  
  ylab("Shannon's Diversity")+
  scale_fill_manual(values = c("Rodents Present" = "white", "Rodents Excluded" = "gray35"))+
  scale_color_manual(values = c("Rodents Present" = "black", "Rodents Excluded" = "black"))  # Black outlines

g2

ggarrange(g1,g2,
          ncol = 2, nrow = 1,
          common.legend = TRUE,
          legend = 'right',
          align = "hv")


```

Presentation Plots

```{r}

ggplot(data = mcc, aes(x = Year, y = Total.Cover, fill = Fire, shape = Fire, colour = Fire)) +
  stat_summary(geom = "point", fun = 'mean', size = 3) +
  stat_summary(geom = "line", fun = "mean", linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Total Plant Cover %") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=16, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=16, family = "serif"),
        axis.title.x  = element_text(size=16, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=16, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=16, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=16, family = "serif", colour = "black"),
        legend.text = element_text(size=16, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
      scale_fill_manual(values = c("Unburned" = "blue", "Burned" = "red")) +
      scale_color_manual(values = c("Unburned" = "blue", "Burned" = "red")) 

ggplot(data = mojdata, aes(x = Year, y = shan_div, fill = Fire, shape = Fire, colour = Fire)) +
  stat_summary(geom = "point", fun = 'mean',  size = 3) +
  stat_summary(geom = "line", fun = "mean",  linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(breaks = seq(min(gbdata$Year), max(gbdata$Year), by = 2)) +  # Set breaks to every other year
  ylab("Shannon's Diversity") +
  xlab(element_blank()) +
  theme(plot.title    = element_text(size=16, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=16, family = "serif"),
        axis.title.x  = element_text(size=16, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=16, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=16, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=16, family = "serif", colour = "black"),
        legend.text = element_text(size=16, family = "serif", colour = "black"),
        legend.title = element_blank(),
        panel.border = element_blank()) +
      scale_fill_manual(values = c("Unburned" = "blue", "Burned" = "red")) +
      scale_color_manual(values = c("Unburned" = "blue", "Burned" = "red")) 

```



Run some stats

```{r}

anocover <- aov(data = mcc, Total.Cover ~ Fire * Rodents)

summary(anocover)

anodiv <- aov(data = mojdata, shan_div ~ Fire * Rodents)

summary(anodiv)

```


Percentage changes

```{r}

# Richness
aggregate(rich ~ Fire, data = mojdata, mean)
# Unburned	6.488889	
# Burned	6.901944
 (6.901944 - 6.488889)/6.488889*100

#Richness increased by 6.365574

# Simpson's diversity
# Percent change for paper
aggregate(simp_div ~ Fire, data = mojdata, mean)
# Unburned	1.737617		
# Burned	1.853886	
 (1.853886 - 1.737617	)/1.737617	*100

# Simpson's diveersity increased by 6.69129

# Shannon's diversity
aggregate(shan_div ~ Fire, data = mojdata, mean)
# Unburned 0.7184890	
# Burned 0.7953773
(0.7953773 - 0.7184890)/0.7184890	*100

#Shannon's diversity increased by 10.70139

# Cover
aggregate(Total.Cover ~ Fire, data = mcc, mean)
# Unburned 60.84322
# Burned 67.03767
67.03767 - 60.84322
# Cover increased by 6.19445

```


I need to create a file specifically with red brome and cheat grass for my rodent structural equation modeling

```{r}

head(gbdata)

cheatgrass <- gbdata |> 
  group_by(Year, Block, Plot) |>
  summarise(BRTE.density = mean(BRTE), TAP = mean(TAP), MAT = mean(MAT))



write.csv(cheatgrass, "BRTE_density.csv")

brome <- mojdata |>
  group_by(Block, Plot, Year) |> 
  summarise(BRRU.density = mean(BRRU), TAP = mean(TAP), MAT = mean(MAT))

write.csv(brome, "BRRU_density.csv")

```
```{r}


mojdata$density <- rowSums(mojdata[, 5:63])
out1 <- mojdata |> select(Year, Block, Plot, BRRU, SCAR, rich, shan_div, simp_div, density, TAP, MAT,
                          Rodent.Abundance, Rodent.Richness, Rodent.Simp, Rodent.Shan, Fire, Rodents)

out2 <- mcc |> select(Year, Block, Plot, Total.Cover)

mojsemdata <- merge(out1, out2, by = c("Year", "Block", "Plot"))



```


Send mccplot data out for SEM in CH 2



```{r}

write.csv(mojsemdata, "MojRodentSEM.csv")

```

Now lets start my model selection

I think I want 15 models that need to be a priori, all should include mixed effects, they are likely to be zero inflated and I need an autoregressive term
glmmTMB should be able to do all of that I think but I need to figure out how.

My first test run will be with richness in the Great Basin because its not zero inflated I don't think

I am going to do a random forest to get relative variable importance to see what should be included in our model

```{r}

library(randomForest)
colnames(gbdata)

# I should scale my data before I do any modeling

### Imma keep an unscaled version for later

gbplotdata <- gbdata
mojplotdata <- mojdata



gbdata[5:31] <- scale(gbdata[c(5:31)])
gbdata[35:58] <- scale(gbdata[35:58])

gbdata <- gbdata |>
  replace(is.na(gbdata), 0)

mojdata[5:62] <- scale(mojdata[5:62])
mojdata[66:89] <- scale(mojdata[66:89])

mojdata <- mojdata |> 
  replace(is.na(mojdata), 0)

colSums(is.na(gbdata))


### GB variable importance
gb.rf.rich <- randomForest(data = gbdata, rich ~ Year + Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents + BRTE + ALAL + CETE +
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )
str(gbdata$rich)
gb.rf.rich
a1 = varImpPlot(gb.rf.rich) # Spring and summmer temp most important, fall rain most important, rodent abundance most important

gb.rf.simp <- randomForest(data = gbdata, simp_div ~ Year + Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents + BRTE+   ALAL + CETE +
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )
a2 = varImpPlot(gb.rf.simp) # Summer Temp, Fall rain, Rodent abundance

gb.rf.shan <- randomForest(data = gbdata, shan_div ~ Year   + Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents + BRTE  + ALAL + CETE +
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )
a3 = varImpPlot(gb.rf.shan)# Summer Temp, Fall rain, Rodent abundance

gb.rf.brte <- randomForest(data = gbdata, BRTE ~ Year  + Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents+
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )
a4 = varImpPlot(gb.rf.brte) # Fire then rodent presence, rodent abundance then winter temp then rodent richness, then fall precip


# Mojave

moj.rf.rich <- randomForest(data = mojdata, rich ~ Year+ Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents+ BRRU + SCAR + 
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )

moj.rf.rich
b1 = varImpPlot(moj.rf.rich) # Fall PPt then spring and summer temp, abundance highest for rodents

moj.rf.simp <- randomForest(data = mojdata, simp_div ~ Year + Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents+ BRRU + SCAR +
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )
b2 = varImpPlot(moj.rf.simp) # spring temp rodent abundance spring and summer ppt

moj.rf.shan <- randomForest(data = mojdata, shan_div ~ Year  + Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents+ BRRU + SCAR +
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )
b3 = varImpPlot(moj.rf.shan) # spring temp spring and summer ppt abundance highest for rodents

moj.rf.brru <- randomForest(data = mojdata, BRRU ~  Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents+
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )
b4= varImpPlot(moj.rf.brru) # TAP, Fall temp, rodent abundance

moj.rf.cover <- randomForest(data = mcc, Total.Cover ~  Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents+
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter )

varImpPlot(moj.rf.cover)

moj.rf.density <- randomForest(data = mojdata, totaldensity ~ Year + Rodent.Abundance + Rodent.Richness + Rodent.Simp + Rodent.Shan + TAP + MAT +
                             SP_Spring +SP_Summer + SP_Fall + SP_Winter + MST_Spring + MST_Summer + MST_Fall + MST_Winter + Fire + Rodents+  
                             lag.TAP + lag.MAT + lag.SP_Spring + lag.SP_Summer + lag.SP_Fall + lag.SP_Winter + lag.MST_Spring + lag.MST_Summer + lag.MST_Fall + lag.MST_Winter)

moj.rf.density

varImpPlot(moj.rf.density)

```

Candidate list of models for GB

```{r}

null <- ' 1+ (1|Block)+ar1(Block + 0 | Year)'

exp <- 'Fire * Rodents + (1|Block)+ ar1(Block + 0 | Year)'

inv <- 'BRTE+ ALAL + CETE +(1|Block) + ar1(Block + 0 | Year)'

tot.clim <- 'TAP + MAT +(1|Block) + ar1(Block + 0 | Year)'

global <- 'TAP + MAT + BRTE + Fire + Rodents + CETE + ALAL +(1|Block) + ar1(Block + 0 | Year)'

rfbest <- 'BRTE + MST_Summer + lag.TAP + ALAL+(1|Block) + ar1(Block + 0 | Year)'

invclim <- 'BRTE + CETE + ALAL + TAP + MAT +(1|Block) + ar1(Block + 0 | Year)'

invlagclim <- 'BRTE + CETE + ALAL + lag.TAP + lag.MAT +(1|Block) + ar1(Block + 0 | Year)'

climrfbest <- 'MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall+(1|Block) + ar1(Block + 0 | Year)'

rfclimfire <- 'MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall + Fire+(1|Block) + ar1(Block + 0 | Year)'

rfclimrod <- 'MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall + Rodents+(1|Block) + ar1(Block + 0 | Year)'

rodents <- 'Rodent.Abundance+(1|Block) + ar1(Block + 0 | Year)'

rodentfire <- 'Rodent.Abundance +Fire +(1|Block) + ar1(Block + 0 | Year)'

rodfireclim <- 'Rodent.Abundance * Fire + Fire*MST_Summer + Fire * lag.TAP +(1|Block) + ar1(Block + 0 | Year)'


```



Great Basin models

Richness

```{r}

hist(gbdata$rich) #nice normal distribution for my poisson data

#lets look for temporal autocorrelation
acf(gbdata$rich)

# These models are going to be defined based on a priori hypotheses

# Null model to compare to as a baseline 
gb.rich.null <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", null))) # Adding AR(1) correlation structure
summary(gb.rich.null)

# Experimental model that considers the treatments we imposed
gb.rich.exp <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", exp))) # Adding AR(1) correlation structure
summary(gb.rich.exp)

# Invasive species model looking at cheatgrass effects
gb.rich.inv <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", inv)))
summary(gb.rich.inv)

# Climate model
gb.rich.clim <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", tot.clim)))
summary(gb.rich.clim)

# Global
gb.rich.global <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", global)))
summary(gb.rich.global)

# RF Best
gb.rich.rfbest <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", rfbest)))
summary(gb.rich.rfbest)


gb.rich.invclim <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", invclim)))

gb.rich.invlagclim <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", invlagclim)))

gb.rich.climrfbest <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", climrfbest)))

gb.rich.rfclimfire <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", rfclimfire)))

gb.rich.rfclimrod <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", rfclimrod)))

gb.rich.rodents <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", rodents)))

gb.rich.rodentsfire <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", rodentfire)))

gb.rich.rodfireclim <- glmmTMB(data = gbdata, 
                        formula = as.formula(paste("rich ~", rodfireclim)))


gb.rich.ints <- glmmTMB(data = gbdata, 
                        rich ~ MST_Summer * Fire + lag.TAP * Fire + MST_Spring * Fire + lag.SP_Fall *Fire +(1|Block)+ ar1(Block + 0 | Year))

gb.rich.intsall <- glmmTMB(data = gbdata, 
                        rich ~ MST_Summer * Fire + lag.TAP * Fire + MST_Spring * Fire + lag.SP_Fall *Fire + Rodent.Abundance*Fire+(1|Block)+ ar1(Block + 0 | Year))

gb.rich.sel <- model.sel(gb.rich.clim, gb.rich.exp, gb.rich.global, gb.rich.inv, gb.rich.invclim, gb.rich.null, 
                         gb.rich.rfbest, gb.rich.rfclimfire, gb.rich.rfclimrod, gb.rich.invlagclim, gb.rich.rodents,
                         gb.rich.rodentsfire, gb.rich.ints, gb.rich.rodfireclim, gb.rich.intsall, gb.rich.climrfbest, gb.rich.null)
gb.rich.sel

```

Lets look at some outputs for the top model

```{r}

summary(gb.rich.ints)

r2(gb.rich.ints)

```
Lets plot those effects

```{r}


figgra<- ggplot(data = gbplotdata, aes(x= MST_Summer, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", aes(fill = Fire), alpha = 0.2) +  # Regression line with CI
  labs(x = "MST Summer (C) ", y = "Richness", colour = "Fire", fill = "Fire") +
  scale_color_manual(values = c("turquoise2", "firebrick2"))+
  scale_fill_manual(values = c("turquoise2", "firebrick2"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme
  
figgra

figgrb<- ggplot(data = gbplotdata, aes(x= lag.TAP, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", aes(fill = Fire), alpha = 0.2) +  # Regression line with CI
  labs(x = "lag TAP (mm)", y = "Richness", colour = "Fire", fill = "Fire") +
   scale_color_manual(values = c("turquoise2", "firebrick2"))+
  scale_fill_manual(values = c("turquoise2", "firebrick2"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgrc<- ggplot(data = gbplotdata, aes(x= MST_Spring, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", aes(fill = Fire), alpha = 0.2) +  # Regression line with CI
  labs(x = "MST Spring (C)", y = "Richness", colour = "Fire", fill = "Fire") +
   scale_color_manual(values = c("turquoise2", "firebrick2"))+
  scale_fill_manual(values = c("turquoise2", "firebrick2"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgrd <- ggplot(data = gbplotdata, aes(x= lag.SP_Fall, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", aes(fill = Fire), alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Fall PPT (mm)", y = "Richness", colour = "Fire", fill = "Fire") +
   scale_color_manual(values = c("turquoise2", "firebrick2"))+
  scale_fill_manual(values = c("turquoise2", "firebrick2"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

ggarrange(figgra, figgrb,
          figgrc, figgrd,
          ncol = 1, nrow = 4,
          common.legend = TRUE,
          legend = "right",
          labels = "AUTO", font.label = list(family="serif", size = 12))



```


```{r}


hist(gbdata$simp_div) #nice normal distribution for my poisson data
summary(gbdata$simp_div)
#lets look for temporal autocorrelation
acf(gbdata$simp_div)

# These models are going to be defined based on a priori hypotheses

# Null model to compare to as a baseline 
gb.simp_div.null <- glmmTMB(data = gbdata,
                            simp_div ~ 1 + (1|Block),
                        family = gaussian()) # Adding AR(1) correlation structure
summary(gb.simp_div.null)

# Experimental model that considers the treatments we imposed
gb.simp_div.exp <- glmmTMB(data = gbdata, 
                        simp_div ~Fire * Rodents + (1|Block)) # Adding AR(1) correlation structure
summary(gb.simp_div.exp)

# Invasive species model looking at cheatgrass effects
gb.simp_div.inv <- glmmTMB(data = gbdata, 
                        simp_div ~ BRTE + CETE + ALAL + (1|Block))
summary(gb.simp_div.inv)

# Climate model
gb.simp_div.clim <- glmmTMB(data = gbdata, 
                        simp_div ~ TAP + MAT +(1|Block))
summary(gb.simp_div.clim)

# Global
gb.simp_div.global <- glmmTMB(data = gbdata, 
                        simp_div ~ TAP + MAT + BRTE + Fire + Rodents + CETE + ALAL + (1|Block))
summary(gb.simp_div.global)

# RF Best
gb.simp_div.rfbest <- glmmTMB(data = gbdata, 
                        simp_div ~ BRTE + MST_Summer + lag.TAP + ALAL+(1|Block))
summary(gb.simp_div.rfbest)


gb.simp_div.invclim <- glmmTMB(data = gbdata, 
                        simp_div ~ BRTE + CETE + ALAL + TAP + MAT +(1|Block))

gb.simp_div.invlagclim <- glmmTMB(data = gbdata, 
                        simp_div ~ BRTE + CETE + ALAL + lag.TAP + lag.MAT +(1|Block))

gb.simp_div.climrfbest <- glmmTMB(data = gbdata, 
                        simp_div ~ MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall+(1|Block))

gb.simp_div.rfclimfire <- glmmTMB(data = gbdata, 
                        simp_div ~ MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall + Fire+(1|Block))

gb.simp_div.rfclimrod <- glmmTMB(data = gbdata, 
                        simp_div~ MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall + Rodents+(1|Block))

gb.simp_div.rodents <- glmmTMB(data = gbdata, 
                        simp_div ~ Rodent.Abundance+(1|Block))

gb.simp_div.rodentsfire <- glmmTMB(data = gbdata, 
                        simp_div ~Rodent.Abundance +Fire +(1|Block) )

gb.simp_div.ints <- glmmTMB(data = gbdata, 
                        simp_div ~ MST_Summer * Fire + lag.TAP * Fire + MST_Spring * Fire + lag.SP_Fall *Fire +(1|Block))

gb.simp_div.intsall <- glmmTMB(data = gbdata, 
                        simp_div ~ MST_Summer * Fire + lag.TAP * Fire + MST_Spring * Fire + lag.SP_Fall *Fire + Rodent.Abundance*Fire+(1|Block))


gb.simp_div.sel <- model.sel(gb.simp_div.clim, gb.simp_div.exp, gb.simp_div.global, gb.simp_div.inv, gb.simp_div.invclim, gb.simp_div.null, 
                         gb.simp_div.rfbest, gb.simp_div.rfclimfire, gb.simp_div.rfclimrod, gb.simp_div.invlagclim, gb.simp_div.rodents,
                         gb.simp_div.rodentsfire, gb.simp_div.climrfbest, gb.simp_div.ints, gb.simp_div.intsall)
gb.simp_div.sel
r2(gb.simp_div.rfclimfire)

```
Lets do Simpsons diversity for the Great Basin

```{r}

summary(gb.simp_div.rfclimfire)


figgia <- ggplot(data = gbplotdata, aes(x= MST_Summer, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "MST Summer (C)", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgib<- ggplot(data = gbplotdata, aes(x= lag.TAP, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag TAP (mm)", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgic<- ggplot(data = gbplotdata, aes(x= MST_Spring, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "MST Spring (C)", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgid<- ggplot(data = gbplotdata, aes(x= lag.SP_Fall, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Fall PPT (mm)", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgie<- ggplot(data = gbplotdata, aes(x= Fire, y = simp_div, fill  = Fire))+
  geom_boxplot()+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  labs(x = "Fire", y = "Simpson's Diversity", colour = "Fire", fill = "Fire") +
    scale_fill_manual(values = c("turquoise2", "firebrick2"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgie

ggarrange(figgia, figgib,
          figgic, figgid,
          figgie,
          ncol = 1, nrow = 4,
          legend = "right",
          labels = "AUTO", font.label = list(family="serif", size = 12))



```


Shannons Diversity


```{r}

# Null model to compare to as a baseline 
gb.shan_div.null <- glmmTMB(data = gbdata,
                            shan_div ~ 1 + (1|Block),
                        family = gaussian()) # Adding AR(1) correlation structure
summary(gb.shan_div.null)

# Experimental model that considers the treatments we imposed
gb.shan_div.exp <- glmmTMB(data = gbdata, 
                        shan_div ~Fire * Rodents + (1|Block)) # Adding AR(1) correlation structure
summary(gb.shan_div.exp)

# Invasive species model looking at cheatgrass effects
gb.shan_div.inv <- glmmTMB(data = gbdata, 
                        shan_div ~ BRTE + CETE + ALAL + (1|Block))
summary(gb.shan_div.inv)

# Climate model
gb.shan_div.clim <- glmmTMB(data = gbdata, 
                        shan_div ~ TAP + MAT +(1|Block))
summary(gb.shan_div.clim)

# Global
gb.shan_div.global <- glmmTMB(data = gbdata, 
                        shan_div ~ TAP + MAT + BRTE + Fire + Rodents + CETE + ALAL + (1|Block))
summary(gb.shan_div.global)

# RF Best
gb.shan_div.rfbest <- glmmTMB(data = gbdata, 
                        shan_div ~ BRTE + MST_Summer + lag.TAP + ALAL+(1|Block))
summary(gb.shan_div.rfbest)


gb.shan_div.invclim <- glmmTMB(data = gbdata, 
                        shan_div ~ BRTE + CETE + ALAL + TAP + MAT +(1|Block))

gb.shan_div.invlagclim <- glmmTMB(data = gbdata, 
                        shan_div ~ BRTE + CETE + ALAL + lag.TAP + lag.MAT +(1|Block))

gb.shan_div.climrfbest <- glmmTMB(data = gbdata, 
                        shan_div ~ MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall+(1|Block))

gb.shan_div.rfclimfire <- glmmTMB(data = gbdata, 
                        shan_div ~ MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall + Fire+(1|Block))

gb.shan_div.rfclimrod <- glmmTMB(data = gbdata, 
                        shan_div~ MST_Summer + lag.TAP + MST_Spring + lag.SP_Fall + Rodents+(1|Block))

gb.shan_div.rodents <- glmmTMB(data = gbdata, 
                        shan_div ~ Rodent.Abundance+(1|Block))

gb.shan_div.rodentsfire <- glmmTMB(data = gbdata, 
                        shan_div ~Rodent.Abundance +Fire +(1|Block) )

gb.shan_div.ints <- glmmTMB(data = gbdata, 
                        shan_div ~ MST_Summer * Fire + lag.TAP * Fire + MST_Spring * Fire + lag.SP_Fall *Fire +(1|Block))

gb.shan_div.intsall <- glmmTMB(data = gbdata, 
                        shan_div ~ MST_Summer * Fire + lag.TAP * Fire + MST_Spring * Fire + lag.SP_Fall *Fire + Rodent.Abundance*Fire+(1|Block))


gb.shan_div.sel <- model.sel(gb.shan_div.clim, gb.shan_div.exp, gb.shan_div.global, gb.shan_div.inv, gb.shan_div.invclim, gb.shan_div.null, 
                         gb.shan_div.rfbest, gb.shan_div.rfclimfire, gb.shan_div.rfclimrod, gb.shan_div.invlagclim, gb.shan_div.rodents,
                         gb.shan_div.rodentsfire, gb.shan_div.climrfbest, gb.shan_div.ints, gb.shan_div.intsall)
gb.shan_div.sel
r2(gb.shan_div.rfclimfire)



```

```{r}

summary(gb.shan_div.rfclimfire)


figgha <- ggplot(data = gbplotdata, aes(x= MST_Summer, y = shan_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "MST Summer (C)", y = "Shannon's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figghb<- ggplot(data = gbplotdata, aes(x= lag.TAP, y = shan_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag TAP (mm)", y = "Shannon's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figghc<- ggplot(data = gbplotdata, aes(x= MST_Spring, y = shan_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "MST Spring (C)", y = "Shannon's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figghd<- ggplot(data = gbplotdata, aes(x= lag.SP_Fall, y = shan_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Fall PPT (mm)", y = "Shannon's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figghe<- ggplot(data = gbplotdata, aes(x= Fire, y = shan_div, fill  = Fire))+
  geom_boxplot()+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  labs(x = "Fire", y = "Shannon's Diversity", colour = "Fire", fill = "Fire") +
    scale_fill_manual(values = c("turquoise2", "firebrick2"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figgie

ggarrange(figgia, figgib,
          figgic, figgid,
          figgie,
          ncol = 1, nrow = 4,
          legend = "right",
          labels = "AUTO", font.label = list(family="serif", size = 12))



```

Now lets do the Mojave


Candidate list of models for Moj

```{r}

null <- ' 1+ar1(Block + 0 | Year)'

exp <- 'Fire * Rodents + ar1(Block + 0 | Year)'

inv <- 'BRRU+ SCAR  + ar1(Block + 0 | Year)'

tot.clim <- 'TAP + MAT  + ar1(Block + 0 | Year)'

global <- 'TAP + MAT + BRRU + Fire + Rodents + SCAR  + ar1(Block + 0 | Year)'

globallag <- 'lag.TAP + lag.MAT + BRRU + Fire + Rodents + SCAR  + ar1(Block + 0 | Year)'

rfbest <- 'BRRU + lag.SP_Spring + lag.TAP + SCAR  + ar1(Block + 0 | Year)'

invclim <- 'BRRU + SCAR + TAP + MAT  + ar1(Block + 0 | Year)'

invlagclim <- 'BRRU  + SCAR + lag.TAP + lag.MAT  + ar1(Block + 0 | Year)'

climrfbest <- 'lag.SP_Spring + lag.TAP + MST_Spring + lag.SP_Winter + ar1(Block + 0 | Year)'

rfclimfire <- 'lag.SP_Spring + lag.TAP + MST_Spring + lag.SP_Winter + Fire + ar1(Block + 0 | Year)'

rfclimrod <- 'lag.SP_Spring + lag.TAP + MST_Spring + lag.SP_Winter + Rodents + ar1(Block + 0 | Year)'

rodents <- 'Rodent.Abundance + ar1(Block + 0 | Year)'

rodentfire <- 'Rodent.Abundance +Fire  + ar1(Block + 0 | Year)'


```



Mojave models

Richness

```{r}

hist(mojdata$rich) #nice normal distribution for my poisson data
varImpPlot(moj.rf.rich)
#lets look for temporal autocorrelation
acf(mojdata$rich)

# These models are going to be defined based on a priori hypotheses



# There is no variance contributed by the random effect
# Conditional model:
#  Groups   Name        Std.Dev.   Corr       
#  Block    (Intercept) 0.00047355            
#  Year     Block1      1.26614167 0.493 (ar1)
#  Residual             0.91760934  

### Level 1 ####

# Null model to compare to as a baseline 
moj.rich.null <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", null))) # Adding AR(1) correlation structure
summary(moj.rich.null)

# Experimental model that considers the treatments we imposed
moj.rich.exp <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", exp))) # Adding AR(1) correlation structure
summary(moj.rich.exp)

# Invasive species model looking at cheatgrass effects
moj.rich.inv <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", inv)))


# Climate model
moj.rich.clim <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", tot.clim)))


# Global
moj.rich.global <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", global)))

moj.rich.globallag <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", globallag)))


moj.rich.invclim <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", invclim)))

moj.rich.invlagclim <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", invlagclim)))

moj.rich.rodents <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", rodents)))

moj.rich.rodentsfire <- glmmTMB(data = mojdata, 
                        formula = as.formula(paste("rich ~", rodentfire)))

#### Level 2 ####

# RF Best
moj.rich.rfbest <- glmmTMB(data = mojdata, 
                                rich ~ lag.TAP + BRRU + SCAR + lag.SP_Spring+
                               ar1(Block + 0 | Year))


moj.rich.climrfbest <- glmmTMB(data = mojdata, 
                              rich ~lag.TAP + lag.SP_Spring + 
                                 ar1(Block + 0 | Year))

moj.rich.rfclimfire <- glmmTMB(data = mojdata, 
                        rich ~lag.TAP * Fire + lag.SP_Spring * Fire + 
                         ar1(Block + 0 | Year))

moj.rich.rfclimrod <- glmmTMB(data = mojdata, 
                       rich ~lag.TAP *Rodent.Abundance +
                          ar1(Block + 0 | Year))


#### Level 3 #####


moj.rich.intsf <- glmmTMB(data = mojdata, 
                        rich ~ lag.TAP * Fire + lag.MAT * Fire + BRRU * Fire + SCAR *Fire 
                        + ar1(Block + 0 | Year))

moj.rich.intsr <- glmmTMB(data = mojdata, 
                        rich ~  lag.TAP * Rodents + lag.MAT * Rodents + BRRU * Rodents + SCAR * Rodents +
                           ar1(Block + 0 | Year))


moj.rich.sel <- model.sel(moj.rich.clim, moj.rich.exp, moj.rich.global, moj.rich.globallag, moj.rich.inv, moj.rich.invclim, moj.rich.null, 
                         moj.rich.rfbest, moj.rich.rfclimfire, moj.rich.rfclimrod, moj.rich.invlagclim, moj.rich.rodents,
                         moj.rich.rodentsfire, moj.rich.climrfbest, moj.rich.intsf, moj.rich.intsr)
moj.rich.sel

moj.rich.sel <- as.data.frame(moj.rich.sel)

write.csv(moj.rich.sel, "Output files/Moj_Rich_Sel.csv")

r2(moj.rich.intsf)
VarCorr(moj.rich.intsf)
summary(moj.rich.intsf)
moj.rich.top <- summary(moj.rich.intsf)
data.frame(moj.rich.top$coefficients$cond)

write.csv(data.frame(moj.rich.top$coefficients$cond), "Output files/Moj_Rich_Top.csv")



```



```{r}

figmra<- ggplot(data = mojplotdata, aes(x= BRRU, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "B. rubens Density", y = "Richness") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = c(1500,1500), label.y = c(2,4)) +
  theme_bw()+
  common_theme


figmrb<- ggplot(data = mojplotdata, aes(x= SCAR, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "S. arabicus Density", y = "Richness") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = c(2500, 2500), label.y = c(3.5,5)) +
  theme_bw()+
  common_theme

figmrc<- ggplot(data = mojplotdata, aes(x= lag.MAT, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Mean Annual Temperature (C)", y = "Richness") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 16.5, label.y = c(3,4)) +
  theme_bw()+ 
  common_theme

figmrd<- ggplot(data = mojplotdata, aes(x= lag.TAP, y = rich, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Total Annual Precipitation (mm)", y = "Richness") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = c(300,300), label.y = c(3,4)) +
  theme_bw()+
  common_theme

ggarrange(figmra, figmrb,
          figmrc, figmrd,
          ncol = 1, nrow = 4,
          legend = "right",
          common.legend = TRUE,
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv",
          heights = 2, widths = 2)

ggarrange(figmra, figmrb,
           figmrd, figmrc,
          ncol = 2, nrow = 2,
          legend = "right",
          common.legend = TRUE,
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv",
          heights = 2, widths = 2)

```


```{r}

# I tested this first with the random effect of block included which resulted in Singularity and block contributed 0 variance
# THese are those results from the top model that included (1|Block):
# Conditional model:
#  Groups   Name        Std.Dev.    Corr       
#  Block    (Intercept) 0.000039948            
#  Year     Block1      0.643989229 0.628 (ar1)
#  Residual             0.377918599  

# So now I am removing the random effect but keeping the autoregressive component

hist(mojdata$simp_div) #nice normal distribution for my poisson data
summary(mojdata$simp_div)
varImpPlot(moj.rf.simp)
#lets look for temporal autocorrelation
acf(mojdata$simp_div)

# These models are going to be defined based on a priori hypotheses

# Null model to compare to as a baseline 
moj.simp_div.null <- glmmTMB(data = mojdata,
                            simp_div ~ 1 + ar1(Block + 0 | Year),
                        family = gaussian()) # Adding AR(1) correlation structure


# Experimental model that considers the treatments we imposed
moj.simp_div.exp <- glmmTMB(data = mojdata, 
                        simp_div ~Fire * Rodents +  ar1(Block + 0 | Year)) # Adding AR(1) correlation structure


# Invasive species model looking at cheatgrass effects
moj.simp_div.inv <- glmmTMB(data = mojdata, 
                        simp_div ~ BRRU  + SCAR +  ar1(Block + 0 | Year))


# Climate model
moj.simp_div.clim <- glmmTMB(data = mojdata, 
                        simp_div ~ TAP + MAT + ar1(Block + 0 | Year))


# Global
moj.simp_div.global <- glmmTMB(data = mojdata, 
                        simp_div ~ TAP + MAT + BRRU + Fire + Rodents + SCAR +  ar1(Block + 0 | Year))

moj.simp_div.globallag <- glmmTMB(data = mojdata, 
                        simp_div ~ lag.TAP + lag.MAT + BRRU + Fire + Rodents + SCAR +  ar1(Block + 0 | Year))

moj.simp_div.invclim <- glmmTMB(data = mojdata, 
                        simp_div ~ BRRU  + SCAR + TAP + MAT + ar1(Block + 0 | Year))

moj.simp_div.invlagclim <- glmmTMB(data = mojdata, 
                        simp_div ~ BRRU + SCAR + lag.TAP + lag.MAT + ar1(Block + 0 | Year))

moj.simp_div.rodents <- glmmTMB(data = mojdata, 
                        simp_div ~ Rodent.Abundance + ar1(Block + 0 | Year))

moj.simp_div.rodentsfire <- glmmTMB(data = mojdata, 
                        simp_div ~Rodent.Abundance +Fire  + ar1(Block + 0 | Year))

### Level 2 ##### 

# RF Best
moj.simp_div.rfbest <- glmmTMB(data = mojdata, 
                        simp_div ~ BRRU + lag.MST_Spring + MST_Winter + SCAR+ ar1(Block + 0 | Year))


moj.simp_div.climrfbest <- glmmTMB(data = mojdata, 
                        simp_div ~ MST_Winter + lag.MST_Spring + ar1(Block + 0 | Year))

moj.simp_div.rfclimfire <- glmmTMB(data = mojdata, 
                        simp_div ~ MST_Winter*Fire + lag.MST_Spring*Fire + ar1(Block + 0 | Year))

moj.simp_div.rfclimrod <- glmmTMB(data = mojdata, 
                        simp_div~ MST_Winter*Rodents + lag.MST_Spring* Rodents+ ar1(Block + 0 | Year))



moj.simp_div.intsf <- glmmTMB(data = mojdata, 
                        simp_div ~ TAP * Fire + MAT * Fire + BRRU  * Fire + Rodents * Fire + SCAR * Fire + 
                         ar1(Block + 0 | Year))

moj.simp_div.intsr <- glmmTMB(data = mojdata, 
                        simp_div ~ TAP * Rodents + MAT * Rodents + BRRU * Rodents +
                          Fire * Rodents + SCAR * Rodents + ar1(Block + 0 | Year))


moj.simp_div.sel <- model.sel(moj.simp_div.clim, moj.simp_div.exp, moj.simp_div.global, moj.simp_div.inv, moj.simp_div.invclim, moj.simp_div.null, moj.simp_div.rfbest, moj.simp_div.rfclimfire, moj.simp_div.rfclimrod, moj.simp_div.invlagclim, moj.simp_div.rodents,
                         moj.simp_div.rodentsfire, moj.simp_div.climrfbest, moj.simp_div.intsr, moj.simp_div.intsf)

moj.simp_div.sel

summary(moj.simp_div.global)
r2(moj.simp_div.global)


performance::check_singularity(moj.simp_div.global)

VarCorr(moj.simp_div.global)

moj.simp_div.sel <- as.data.frame(moj.simp_div.sel)

write.csv(moj.simp_div.sel, "Output files/Moj_Simp_Sel.csv")

moj.simp.top <- summary(moj.simp_div.global)

write.csv(data.frame(moj.simp.top$coefficients$cond), "Output files/Moj_Simp_Top.csv")


```

Mojave Simpsons Diversity

```{r}

moj.simp_div.sel
summary(moj.simp_div.global)

figmia <- ggplot(data = mojplotdata, aes(x= BRRU, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  geom_smooth(method = "lm",alpha = 0.2) +  # Regression line with CI
  labs(x = "BRRU Density", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme


figmib <- ggplot(data = mojplotdata, aes(x= SCAR, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "SCAR Density", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figmic<- ggplot(data = mojplotdata, aes(x= MAT, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "MAT (C)", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figmid<- ggplot(data = mojplotdata, aes(x= TAP, y = simp_div))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "TAP (mm)", y = "Simpson's Diversity") +
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

figmie <- ggplot(data = mojplotdata, aes(x= Fire, y = simp_div, fill  = Fire))+
  geom_boxplot()+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  labs(x = "Fire", y = "Simpson's Diversity", colour = "Fire", fill = "Fire") +
    scale_fill_manual(values = c("turquoise2", "firebrick2"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme+
  theme(legend.position = c(0.5, 0.7),  # Adjust position (x, y)
        legend.background = element_rect(fill = "white", color = "black"),  # Background box
        legend.key = element_rect(fill = "white"))+ 
  theme(legend.background = element_rect(color = NA))

figmif <- ggplot(data = mojplotdata, aes(x= Rodents, y = simp_div, fill  = Rodents))+
  geom_boxplot()+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  labs(x = "Rodents", y = "Simpson's Diversity", colour = "Rodents", fill = "Rodents") +
    scale_fill_manual(values = c("yellow", "purple" ), labels= c("Present", "Excluded"))+
  #stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme+
  theme(legend.position = c(0.5, 0.18),  # Adjust position (x, y)
        legend.background = element_rect(fill = "white", color = "black"),  # Background box
        legend.key = element_rect(fill = "white"))+ 
  theme(legend.background = element_rect(color = NA))

ggarrange(figmia, figmib,
          figmic, figmid,
          figmie, figmif,
          ncol = 2, nrow = 3,
          labels = "AUTO", font.label = list(family="serif", size = 12))

```



Shannons Diversity


```{r}

hist(mojdata$shan_div) #nice normal distribution for my poisson data
summary(mojdata$shan_div)
#lets look for temporal autocorrelation
acf(mojdata$shan_div)
varImpPlot(moj.rf.shan)
# These models are going to be defined based on a priori hypotheses

# Null model to compare to as a baseline 
moj.shan_div.null <- glmmTMB(data = mojdata,
                            shan_div ~ 1 + (1|Block)+ ar1(Block + 0 | Year),
                        family = gaussian()) # Adding AR(1) correlation structure
summary(moj.shan_div.null)

# Experimental model that considers the treatments we imposed
moj.shan_div.exp <- glmmTMB(data = mojdata, 
                        shan_div ~Fire * Rodents + (1|Block)+ ar1(Block + 0 | Year)) # Adding AR(1) correlation structure
summary(moj.shan_div.exp)

# Invasive species model looking at cheatgrass effects
moj.shan_div.inv <- glmmTMB(data = mojdata, 
                        shan_div ~ BRRU +  SCAR + (1|Block)+ ar1(Block + 0 | Year))
summary(moj.shan_div.inv)

# Climate model
moj.shan_div.clim <- glmmTMB(data = mojdata, 
                        shan_div ~ TAP + MAT +(1|Block)+ ar1(Block + 0 | Year))
summary(moj.shan_div.clim)

# Global
moj.shan_div.global <- glmmTMB(data = mojdata, 
                        shan_div ~ TAP + MAT + BRRU + Fire + Rodents + SCAR + (1|Block)+ ar1(Block + 0 | Year))

# Global
moj.shan_div.globallag <- glmmTMB(data = mojdata, 
                        shan_div ~ lag.TAP + lag.MAT + BRRU + Fire + Rodents + SCAR + (1|Block)+ ar1(Block + 0 | Year))

moj.shan_div.invclim <- glmmTMB(data = mojdata, 
                        shan_div ~ BRRU + SCAR + TAP + MAT +(1|Block)+ ar1(Block + 0 | Year))

moj.shan_div.invlagclim <- glmmTMB(data = mojdata, 
                        shan_div ~ BRRU + SCAR + lag.TAP + lag.MAT +(1|Block)+ ar1(Block + 0 | Year))

moj.shan_div.rodents <- glmmTMB(data = mojdata, 
                        shan_div ~ Rodent.Abundance+(1|Block)+ ar1(Block + 0 | Year))

moj.shan_div.rodentsfire <- glmmTMB(data = mojdata, 
                        shan_div ~Rodent.Abundance +Fire +(1|Block) + ar1(Block + 0 | Year))

### Level 2 ####
moj.shan_div.rfbest <- glmmTMB(data = mojdata, 
                        shan_div ~ BRRU + lag.SP_Winter + MST_Winter + SCAR+(1|Block)+ ar1(Block + 0 | Year))



moj.shan_div.climrfbest <- glmmTMB(data = mojdata, 
                        shan_div ~ lag.SP_Winter + MST_Winter + (1|Block)+ ar1(Block + 0 | Year))

moj.shan_div.rfclimfire <- glmmTMB(data = mojdata, 
                        shan_div ~ lag.SP_Winter * Fire + MST_Winter * Fire + (1|Block)+ ar1(Block + 0 | Year))

moj.shan_div.rfclimrod <- glmmTMB(data = mojdata, 
                        shan_div~ lag.SP_Winter * Rodents + MST_Winter * Rodents + (1|Block)+ ar1(Block + 0 | Year))



moj.shan_div.intsf <- glmmTMB(data = mojdata, 
                        shan_div ~ lag.TAP * Fire + lag.MAT * Fire + BRRU  * Fire  + SCAR * Fire +
                        (1|Block)+ ar1(Block + 0 | Year))

moj.shan_div.intsr <- glmmTMB(data = mojdata, 
                        shan_div ~ lag.TAP * Rodents + lag.MAT * Rodents + BRRU * Rodents +
                         SCAR * Rodents +(1|Block)+ ar1(Block + 0 | Year))


moj.shan_div.sel <- model.sel(moj.shan_div.clim, moj.shan_div.exp, moj.shan_div.global,moj.shan_div.globallag, moj.shan_div.inv, moj.shan_div.invclim, moj.shan_div.null, moj.shan_div.rfbest, moj.shan_div.rfclimfire, moj.shan_div.rfclimrod, moj.shan_div.invlagclim, moj.shan_div.rodents,
                         moj.shan_div.rodentsfire, moj.shan_div.climrfbest, moj.shan_div.intsf, moj.shan_div.intsr)

moj.shan_div.sel

summary(moj.shan_div.intsf)

r2(moj.shan_div.intsf)

performance::check_singularity(moj.shan_div.intsf) # Not singular


moj.shan_div.sel <- as.data.frame(moj.shan_div.sel)

write.csv(moj.shan_div.sel, "Output files/Moj_Shan_Sel.csv")

moj.shan.top <- summary(moj.shan_div.intsf)

write.csv(data.frame(moj.shan.top$coefficients$cond), "Output files/Moj_Shan_Top.csv")

```


I want to graph the responses of the variables from the top models but not using the predictions



```{r}


figmha<- ggplot(data = mojplotdata, aes(x= BRRU, y = shan_div, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "B. rubens Density", y = "Shannon's Diveristy") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = c(1500,1500), label.y = c(0.3,0.5)) +
  theme_bw()+
  common_theme


figmhb<- ggplot(data = mojplotdata, aes(x= SCAR, y = shan_div, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  #stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "S. arabicus Density", y = "Shannon's Diversity") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = c(2500, 2500), label.y = c(1,1.3)) +
  theme_bw()+
  common_theme

figmhc<- ggplot(data = mojplotdata, aes(x= lag.MAT, y = shan_div, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Mean Annual Temperature (C)", y = "Shannon's Diversity") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 16.5, label.y = c(0.3,0.5)) +
  theme_bw()+ 
  common_theme

figmhd<- ggplot(data = mojplotdata, aes(x= lag.TAP, y = shan_div, colour  = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Total Annual Precipitation (mm)", y = "Shannon's Diversity") +
  scale_color_manual(values = c("turquoise2", "firebrick2"), labels = c("Unburned", "Burned"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = c(300,300), label.y = c(0.3,0.5)) +
  theme_bw()+
  common_theme

ggarrange(figmha, figmhb,
          figmhc, figmhd,
          ncol = 1, nrow = 4,
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv",
          heights = 2, widths = 2,
          common.legend = TRUE,
          legend = "right")

ggarrange(figmha, figmhb,
           figmhd, figmhc,
          ncol = 2, nrow = 2,
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv",
          heights = 2, widths = 2,
          common.legend = TRUE,
          legend = "right")

figmhd

```

Cover

```{r}

mccplotdata <- mcc
# I need to standardize mcc
mcc[6:63] <- scale(mcc[6:63])
mcc[67:90] <- scale(mcc[67:90])



# I am removing the autoregressive component because of the small variance associated with year leading to model convergence
# Here are the results that say that
# VarCorr(moj.Total.Cover.inv)
# 
# Conditional model:
#  Groups   Name        Std.Dev.   Corr       
#  Block    (Intercept)  2.0507467            
#  Year     Block1       0.0056366 0.986 (ar1)
#  Residual             13.4689436  


str(mcc)
str(mojdata)
hist(mcc$Total.Cover) #nice normal distribution for my poisson data
summary(mcc$Total.Cover)

varImpPlot(moj.rf.cover)

#lets look for temporal autocorrelation
acf(mcc$Total.Cover)
mcc$Block <- as.factor(mcc$Block)
# These models are going to be defined based on a priori hypotheses

# Null model to compare to as a baseline 
moj.Total.Cover.null <- glmmTMB(data = mcc,
                            Total.Cover ~ 1 + (1|Block),
                        family = gaussian()) # Adding AR(1) correlation structure
summary(moj.Total.Cover.null)

# Experimental model that considers the treatments we imposed
moj.Total.Cover.exp <- glmmTMB(data = mcc, 
                        Total.Cover ~Fire * Rodents + (1|Block)) # Adding AR(1) correlation structure
summary(moj.Total.Cover.exp)

# Invasive species model looking at cheatgrass effects
moj.Total.Cover.inv <- glmmTMB(data = mcc, 
                        Total.Cover ~ BRRU +  SCAR + (1|Block))
summary(moj.Total.Cover.inv)

# Climate model
moj.Total.Cover.clim <- glmmTMB(data = mcc, 
                        Total.Cover ~ TAP + MAT +(1|Block))
summary(moj.Total.Cover.clim)

# Global
moj.Total.Cover.global <- glmmTMB(data = mcc, 
                        Total.Cover ~ TAP + MAT + BRRU + Fire + Rodents + SCAR + (1|Block))

moj.Total.Cover.globallag <- glmmTMB(data = mcc, 
                        Total.Cover ~ lag.TAP + lag.MAT + BRRU + Fire + Rodents + SCAR + (1|Block))


moj.Total.Cover.invclim <- glmmTMB(data = mcc, 
                        Total.Cover ~ BRRU + SCAR + TAP + MAT +(1|Block))

moj.Total.Cover.invlagclim <- glmmTMB(data = mcc, 
                        Total.Cover ~ BRRU + SCAR + lag.TAP + lag.MAT +(1|Block))

moj.Total.Cover.rodents <- glmmTMB(data = mcc, 
                        Total.Cover ~ Rodent.Abundance+(1|Block))

moj.Total.Cover.rodentsfire <- glmmTMB(data = mcc, 
                        Total.Cover ~Rodent.Abundance +Fire +(1|Block) )


# Level 2

moj.Total.Cover.rfbest <- glmmTMB(data = mcc, 
                        Total.Cover ~ lag.MAT + lag.MST_Winter + Rodent.Abundance + Fire + (1|Block))


moj.Total.Cover.climrfbest <- glmmTMB(data = mcc, 
                        Total.Cover ~  lag.MAT + lag.MST_Winter + (1|Block))

moj.Total.Cover.rfclimfire <- glmmTMB(data = mcc, 
                        Total.Cover ~ lag.MAT * Fire + lag.MST_Winter * Fire +(1|Block))

moj.Total.Cover.rfclimrod <- glmmTMB(data = mcc, 
                        Total.Cover~ lag.MAT* Rodents + lag.MST_Winter* Rodents +(1|Block))



moj.Total.Cover.intsf <- glmmTMB(data = mcc, 
                        Total.Cover ~ BRRU * Fire + SCAR * Fire + lag.TAP * Fire + lag.MAT * Fire
                        +(1|Block))

moj.Total.Cover.intsr <- glmmTMB(data = mcc, 
                        Total.Cover ~ BRRU * Rodents + SCAR * Rodents + lag.TAP * Rodents + lag.MAT * Rodents +(1|Block))


moj.Total.Cover.sel <- model.sel(moj.Total.Cover.clim, moj.Total.Cover.exp, moj.Total.Cover.global,  moj.Total.Cover.inv, moj.Total.Cover.invclim, moj.Total.Cover.null, 
                         moj.Total.Cover.rfbest, moj.Total.Cover.rfclimfire, moj.Total.Cover.rfclimrod, moj.Total.Cover.invlagclim, moj.Total.Cover.rodents,
                         moj.Total.Cover.rodentsfire, moj.Total.Cover.climrfbest, moj.Total.Cover.intsf, moj.Total.Cover.intsr)
moj.Total.Cover.sel

summary(moj.Total.Cover.intsf)
VarCorr(moj.Total.Cover.intsf)
r2(moj.Total.Cover.intsf)
performance::check_singularity(moj.Total.Cover.intsf)

performance::check_singularity(moj.Total.Cover.intsf) # Not singular

moj.Total.Cover.sel <- as.data.frame(moj.Total.Cover.sel)

write.csv(moj.Total.Cover.sel, "Output files/Moj_Cover_Sel.csv")

moj.cover.top <- summary(moj.Total.Cover.intsf)

write.csv(data.frame(moj.cover.top$coefficients$cond), "Output files/Moj_Cover_Top.csv")

```

```{r}

summary(moj.Total.Cover.intsf)

figmca<- ggplot(data = mccplotdata, aes(x= BRRU, y = Total.Cover, color = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "B. rubens Density", y = "Plant Cover (%)") +
  scale_color_manual(values = c("turquoise2", "firebrick2"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 1500, label.y = c(40,50)) +
  theme_bw()+
  common_theme

figmcb<- ggplot(data = mccplotdata, aes(x= SCAR, y = Total.Cover, color = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "S. arabicus Density", y = "Plant Cover (%)") +
  scale_color_manual(values = c("turquoise2", "firebrick2"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 2500, label.y = c(40,50)) +
  theme_bw()+
  common_theme

figmcc<- ggplot(data = mccplotdata, aes(x= lag.TAP, y = Total.Cover, color = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Total Annual Precipitation (mm)", y = "Plant Cover (%)") +
  scale_color_manual(values = c("turquoise2", "firebrick2"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 100, label.y = c(75,85)) +
  theme_bw()+ 
  common_theme

figmcd<- ggplot(data = mccplotdata, aes(x= lag.MAT, y = Total.Cover, color = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "lag Mean Annual Temperature (C)", y = "Plant Cover (%)") +
  scale_color_manual(values = c("turquoise2", "firebrick2"))+
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 16.25, label.y = c(45,55)) +
  theme_bw()+
  common_theme

ggarrange(figmca, figmcb,
          figmcc, figmcd,
          ncol = 1, nrow = 4,
          legend = "right",
          common.legend = TRUE,
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv",
          heights = 2, widths = 2)

ggarrange(figmca, figmcb,
          figmcc, figmcd,
          ncol = 2, nrow = 2,
          legend = "right",
          common.legend = TRUE,
          labels = "AUTO", font.label = list(family="serif", size = 12),
          align = "hv",
          heights = 2, widths = 2)

figmcb

```


Total Density



```{r}

varImpPlot(moj.rf.density)

#lets look for temporal autocorrelation
acf(mojdata$totaldensity)
mojdata$Block <- as.factor(mojdata$Block)
# These models are going to be defined based on a priori hypotheses

# Null model to compare to as a baseline 
moj.totaldensity.null <- glmmTMB(data = mojdata,
                            totaldensity ~ 1 + ar1(Block + 0 | Year),
                        family = gaussian()) # Adding AR(1) correlation structure
summary(moj.totaldensity.null)

# Experimental model that considers the treatments we imposed
moj.totaldensity.exp <- glmmTMB(data = mojdata, 
                        totaldensity ~Fire * Rodents + ar1(Block + 0 | Year)) # Adding AR(1) correlation structure
summary(moj.totaldensity.exp)


# Climate model
moj.totaldensity.clim <- glmmTMB(data = mojdata, 
                        totaldensity ~ TAP + MAT +ar1(Block + 0 | Year))
summary(moj.totaldensity.clim)

# Global
moj.totaldensity.global <- glmmTMB(data = mojdata, 
                        totaldensity ~ TAP + MAT +  Fire + Rodents  + ar1(Block + 0 | Year))

moj.totaldensity.globallag <- glmmTMB(data = mojdata, 
                        totaldensity ~ lag.TAP + lag.MAT + Fire + Rodents  + ar1(Block + 0 | Year))


moj.totaldensity.rodents <- glmmTMB(data = mojdata, 
                        totaldensity ~ Rodent.Abundance + ar1(Block + 0 | Year))

moj.totaldensity.rodentsfire <- glmmTMB(data = mojdata, 
                        totaldensity ~Rodent.Abundance +Fire + ar1(Block + 0 | Year) )


# Level 2

moj.totaldensity.rfbest <- glmmTMB(data = mojdata, 
                        totaldensity ~ Fire + SP_Fall +MST_Summer + MST_Winter+ ar1(Block + 0 | Year))


moj.totaldensity.climrfbest <- glmmTMB(data = mojdata, 
                        totaldensity ~  SP_Fall+ MST_Winter + ar1(Block + 0 | Year))

moj.totaldensity.rfclimfire <- glmmTMB(data = mojdata, 
                        totaldensity ~ SP_Fall * Fire + MST_Winter * Fire + ar1(Block + 0 | Year))

moj.totaldensity.rfclimrod <- glmmTMB(data = mojdata, 
                        totaldensity~ SP_Fall * Rodents + SP_Fall* Rodents + ar1(Block + 0 | Year))



moj.totaldensity.intsf <- glmmTMB(data = mojdata, 
                        totaldensity ~ BRRU * Fire + SCAR * Fire + lag.TAP * Fire + lag.MAT * Fire
                        +(1|Block))

moj.totaldensity.intsr <- glmmTMB(data = mojdata, 
                        totaldensity ~ BRRU * Rodents + SCAR * Rodents + lag.TAP * Rodents + lag.MAT * Rodents +(1|Block))


moj.totaldensity.sel <- model.sel(moj.totaldensity.clim, moj.totaldensity.exp, moj.totaldensity.global, moj.totaldensity.globallag,   
                      moj.totaldensity.null, moj.totaldensity.rfbest, moj.totaldensity.rfclimfire, moj.totaldensity.rfclimrod, moj.totaldensity.rodents,
                         moj.totaldensity.rodentsfire, moj.totaldensity.climrfbest)

moj.totaldensity.sel

summary(moj.totaldensity.rfclimfire)
VarCorr(moj.totaldensity.rfclimfire)
r2(moj.totaldensity.rfclimfire)
performance::check_singularity(moj.totaldensity.rfclimfire)

performance::check_singularity(moj.totaldensity.rfclimfire) # Not singular

moj.totaldensity.sel <- as.data.frame(moj.totaldensity.sel)

write.csv(moj.totaldensity.sel, "Output files/Moj_Density_Sel.csv")

moj.cover.top <- summary(moj.totaldensity.rfclimfire)

write.csv(data.frame(moj.cover.top$coefficients$cond), "Output files/Moj_Density_Top.csv")

```


```{r}

#summary(moj.totaldensity.rfclimfire)

figmda <- ggplot(data = mojplotdata, aes(x= SP_Fall, y = totaldensity, color = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm",alpha = 0.2) +  # Regression line with CI
  labs(x = "Fall Precipitation (mm)", y = "Total Plant Density") +
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 60, label.y = c(2000, 2200)) +
  scale_color_manual(values = c("turquoise2", "firebrick2"))+
  theme_bw()+
  coord_cartesian(ylim = c(0,2500))+
  common_theme


figmdb <- ggplot(data = mojplotdata, aes(x= MST_Winter, y = totaldensity, color = Fire))+
  stat_summary(geom = "point", fun = "mean")+
  stat_summary(geom = "errorbar")+  # Error bars (mean  SE)
  geom_smooth(method = "lm", alpha = 0.2) +  # Regression line with CI
  labs(x = "Winter Temperature (C)", y = "Total Plant Density") +
  stat_regline_equation(aes(label = paste( ..rr.label..)), label.x = 6, label.y = c(2100, 2300)) +
  scale_color_manual(values = c("turquoise2", "firebrick2"))+
  theme_bw()+
  coord_cartesian(ylim = c(0,2500))+
  common_theme



ggarrange(figmda, figmdb,
          ncol = 1, nrow = 2,
          labels = "AUTO", font.label = list(family="serif", size = 12),
          common.legend = TRUE, 
          legend = "right",
          align = "hv",
          widths = c(2,2))

```


I should test if BRTE decreases cete (Biotic resistance paper)




```{r}

mccplotdata$Total.Density<-rowSums(mccplotdata[7:64])

ggplot(data = mccplotdata, aes(x= Total.Cover, y = Total.Density))+
  stat_summary(geom = "point", fun = "mean")+
  geom_smooth(method = "loess", alpha = 0.2) +  # Regression line with CI
  stat_regline_equation(aes(label = paste(after_stat(eq.label), ..rr.label.., sep = "~~~"))) +
  theme_bw()+
  common_theme

```




```{r}
#Red brome cover
mean(mccplotdata$BRRU)
mean(gbcc$BRTE.cover)

brte.plot
moj.brru.plot
```


I want to look at the interactive effects of fire and rodents with climate

#Richness

```{r}

library(lmerTest)
acf(mojplotdata$rich)
rich.fp <- glmmTMB(data = mojplotdata, rich ~ Fire * lag.TAP +(1|Block) + ar1(Block + 0 | Year))
rich.ft <- glmmTMB(data = mojplotdata, rich ~ Fire * lag.MAT + (1|Block) + ar1(Block + 0 | Year))

rich.rp <- glmmTMB(data = mojplotdata, rich ~ Rodents * lag.TAP + (1|Block) + ar1(Block + 0 | Year))
rich.rt <- glmmTMB(data = mojplotdata, rich ~ Rodents * lag.MAT + (1|Block) + ar1(Block + 0 | Year))

summary(rich.ft)

```
# Density

```{r}

acf(mojplotdata$totaldensity)
hist(log(mojplotdata$totaldensity))
VarCorr(dens.fp)  
dens.fp <- glmmTMB(data = mojdata, totaldensity ~ Fire * lag.TAP  + ar1(Block + 0 | Year))
dens.ft <- glmmTMB(data = mojdata, totaldensity ~ Fire * lag.MAT  + ar1(Block + 0 | Year))

dens.rp <- glmmTMB(data = mojdata, totaldensity ~ Rodents * lag.TAP + ar1(Block + 0 | Year))
dens.rt <- glmmTMB(data = mojdata, totaldensity ~ Rodents * lag.MAT + ar1(Block + 0 | Year))

summary(dens.r)

```

# Shannons Diversity

```{r}

acf(mojdata$shan_div)
hist(mojdata$shan_div)
VarCorr(shan.fp)  
shan.fp <- glmmTMB(data = mojdata, shan_div ~ Fire * lag.TAP  + ar1(Block + 0 | Year))
shan.ft <- glmmTMB(data = mojdata, shan_div ~ Fire * lag.MAT  + ar1(Block + 0 | Year))

shan.rp <- glmmTMB(data = mojdata, shan_div ~ Rodents * lag.TAP + ar1(Block + 0 | Year))
shan.rt <- glmmTMB(data = mojdata, shan_div ~ Rodents * lag.MAT + ar1(Block + 0 | Year))

summary(shan.rp)

```


# Cover

```{r}

acf(mcc$Total.Cover)
hist(mcc$Total.Cover)
VarCorr(cover.fp)  
cover.fp <- glmmTMB(data = mcc, Total.Cover ~ Fire * lag.TAP  + ar1(Block + 0 | Year))
cover.ft <- glmmTMB(data =  mcc, Total.Cover ~ Fire * lag.MAT  + ar1(Block + 0 | Year))

cover.rp <- glmmTMB(data =  mcc, Total.Cover ~ Rodents * lag.TAP + ar1(Block + 0 | Year))
cover.rt <- glmmTMB(data =  mcc, Total.Cover ~ Rodents * lag.MAT + ar1(Block + 0 | Year))

summary(cover.ft)

```
Export models

```{r}

library(broom)
library(dplyr)
library(readr)

# List of your models
models <- list(rich.fp, rich.ft, rich.rp, rich.rt, dens.fp, dens.ft, dens.rp, dens.rt,
               shan.fp, shan.ft, shan.rp, shan.rt, cover.fp, cover.ft, cover.rp, cover.rt)
# Extract fixed effects
fixed_effects <- bind_rows(lapply(seq_along(models), function(i) {
  coef_df <- as.data.frame(summary(models[[i]])$coefficients$cond)  # Extract fixed effects
  coef_df$Parameter <- rownames(coef_df)  # Add parameter names
  coef_df$Model <- paste0("Model_", i)  # Add model identifier
  coef_df
}), .id = "Model_Index")

# Rename columns for clarity
colnames(fixed_effects) <- c("Model_Index", "Estimate", "Std_Error", "z_value", "p_value", "Parameter", "Model")



model_metrics$name <- c('rich.fp', 'rich.ft', 'rich.rp', 'rich.rt', 'dens.fp', 'dens.ft', 'dens.rp', 'dens.rt',
               'shan.fp', 'shan.ft', 'shan.rp', 'shan.rt', 'cover.fp', 'cover.ft', 'cover.rp', 'cover.rt')

# Save to CSV
write_csv(fixed_effects, "model_summaries.csv")



```

