---
title: "Desert Plant Models Great Basin Data Cleaning"
author: "R Pienaar"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Libraries

```{r}

library(tidyr)
library(tidyverse)
library(ggplot2)
library(vegan)

```


The goal of this project is to understand how fire, climate, and rodents affect native and introduced plants in the Great Basin and Mojave Deserts. I will use data collected from 2013 to 2025 from the study plots that includes rodent species occurence and plant density transects. Climate data will be extracted using PRISM. There is no variation between the experimental blocks in terms of climatic differences but we are looking at differences between the 2 deserts.

I am going to use 1 file for both deserts but I will use seperate input data each time I run to get the seperate results for the different sites

I will start by loading all of the Great Basin Files

First the plant data

```{r}

gbplants <- read.csv("Input files/Rush_Veg_Density.csv")

```
    
      I need to format the plant data correctly
      
```{r}

head(gbplants)

# Get rid of the individual quadrats
gbplants <- gbplants |> select(DATE, BLOCK, PLOT, BURNED, RODENTS, TRANSECT, SPECIES, Density..m.2.)

# First thing I wanna do is rename the species column cos calling it plant.species was dumb
colnames(gbplants)
colnames(gbplants) = c("DATE", "Block", "Plot", "Burned", "Rodents", "Transect", "Species", "Density")

### How many unique species do we have
unique(gbplants$Species) ### This is messy we have 73 species here
gbplants$Species = as.factor(gbplants$Species)
summary(gbplants$Species)

# Lets go through em one by one,

# ? We'll just remove
gbplants = subset(gbplants, !(Species %in% c("?", "no living plants", "UK 1", "UK1", "uk2", "UKI", "UKL", "Unknown purple", "Vetch?", "Wheat grass", "rUSh gilia", "ELEL TILLERS")))

# Reset the factor levels
gbplants$Species <- as.character(gbplants$Species)
gbplants$Species <- as.factor(gbplants$Species)

# All the ELELs are the same thing
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("BELEL", "ELEL COUNT", "ELEL5", "ELELS", "ELE", "ELEL 5") ~ "ELEL",
    TRUE ~ Species
  ))

# Now do CETE
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("CETE5", "CETI") ~ "CETE",
    TRUE ~ Species
  ))

# Now ARTR
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("ARTR SEEDLING", "ARTRW", "ARTRW8") ~ "ARTR",
    TRUE ~ Species
  ))

# Now BRTE
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("CBRTE") ~ "BRTE",
    TRUE ~ Species
  ))

gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("LEPE2", "LEPI") ~ "LEPE",
    TRUE ~ Species
  ))

## SIAL
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("SIAL 2", "SIAL2") ~ "SIAL",
    TRUE ~ Species
  ))

# HAGL 
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("HAGL ") ~ "HAGL",
    TRUE ~ Species
  ))

# RATE
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("RTE") ~ "RATE",
    TRUE ~ Species
  ))

# POSA
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("POSA?", "POSA") ~ "POSE",
    TRUE ~ Species
  ))

gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("HUMMING BIRD") ~ "CAMI",
    TRUE ~ Species
  ))

#MEAL
gbplants <- gbplants %>%
  mutate(Species = case_when(
    Species %in% c("MEAL6", "MENTZELIA") ~ "MEAL",
    TRUE ~ Species
  ))

# Remove individual plants only found once
gbplants <- gbplants %>%
  group_by(Species) %>%
  filter(n() > 1) %>%
  ungroup()

# We arent interested in the 0's right now

#gbplants <- gbplants %>%
#  filter(Density != 0)

gbplants$Plot <- as.factor(gbplants$Plot)
gbplants$Burned <- as.factor(gbplants$Burned)
gbplants$Rodents <- as.factor(gbplants$Rodents)
gbplants$Density <- as.numeric(gbplants$Density)


summary(gbplants)

#colnames(gbdata)
#colnames(mojdata)

```
  
Now I think I want to restructure the dataframe so that there is a column for every species

```{r}

head(gbplants)

# Group by the specified columns and count occurrences
unique_test <- gbplants %>%
  group_by(DATE, Block, Plot, Species, Transect) %>%
  summarise(count = n(), .groups = 'drop')

# Filter the results to find any combinations that occur more than once
duplicates <- unique_test %>%
  filter(count > 1)


gbplants <- gbplants %>%
  group_by(DATE, Block, Plot, Transect, Species) %>%
  summarise(Density = mean(Density), .groups = 'drop')

# Reshape the data using pivot_wider
gbplants <- gbplants %>%
  pivot_wider(names_from = Species, values_from = Density, values_fill = 0)

### At this point I have all my plant data in the format that I want it

# Now I am going to calculate richness and diversity

gbplants$rich <- rowSums(gbplants[,5:32]> 0 )
gbplants$simp_div <- diversity(gbplants[,5:32], index = "invsimpson")

summary(gbplants$simp_div)

gbplants$simp_div[gbplants$simp_div==Inf] <- 1

gbplants$shan_div <- diversity(gbplants[,5:32], index = "shannon")

summary(gbplants)

head(gbplants)
```

I am going to reformat this so that I have one row for each plot each year

```{r}

gbdata <- gbplants %>%
  group_by(DATE, Block, Plot) %>%
  summarise(across(HAGL:shan_div, mean, na.rm = TRUE), .groups = 'drop')


# Create a complete dataset
gbdata <- gbdata %>%
  complete(DATE = full_seq(DATE, 1),
           Block = 1:5,
           Plot = factor(c('BN', 'UN', 'BS', 'US')))

sum(is.na(gbdata$Plot))

gbdata <- gbdata |>  mutate_all(~replace(., is.na(.), 0))

summary(gbdata)

```

I am going to reformat this so that I have one row for each plot each year

```{r}

gbdata <- gbplants %>%
  group_by(DATE, Block, Plot) %>%
  summarise(across(ALAL:shan_div, mean, na.rm = TRUE), .groups = 'drop')


# Create a complete dataset
gbdata <- gbdata %>%
  complete(DATE = full_seq(DATE, 1),
           Block = seq(1:5),
           Plot = factor(c('BN', 'UN', 'BS', 'US')))

sum(is.na(gbdata$Plot))

gbdata <- gbdata |>  mutate_all(~replace(., is.na(.), 0))

summary(gbdata)

```

    Next up I want to get the rodent data
  
```{r}

gbrodents <- read.csv("C:/Users/ryanp/Documents/Academia/Dissertation/Chapter2_Rodents/Rodent_analysis/Rush_Rodents_CH2_Output.csv")

head(gbrodents)

gbrodents$Plot <- as.factor(paste0(gbrodents$Fire, gbrodents$Rodents))


gbrodents = gbrodents|> select(Year, BLOCK, Fire, Rodents, abundance, richness, simp_div, shan_div)

gbrodents <- gbrodents|>
  group_by(Year, BLOCK, Fire, Rodents) %>%
  summarise(Rodent.Abundance = mean(abundance), 
            Rodent.Richness = mean(richness), 
            Rodent.Simp = mean(simp_div), 
            Rodent.Shan = mean(shan_div),
            .groups = 'drop')

gbrodents$Plot <- as.factor(paste0(gbrodents$Fire, gbrodents$Rodents))
colnames(gbrodents) = c("DATE", "Block", "Fire", "Rodents", "Rodent.Abundance", "Rodent.Richness", "Rodent.Simp", "Rodent.Shan", "Plot")

head(gbrodents)
summary(gbrodents)
head(gbplants)

gbdata = merge(gbdata, gbrodents[,c("DATE", "Block", "Plot", "Rodent.Abundance", "Rodent.Richness", "Rodent.Simp", "Rodent.Shan")], 
                   by = c("DATE", "Block", "Plot"), all.x = TRUE)

gbdata <- gbdata %>% replace(is.na(.), 0)

head(gbdata)

```

    Now I need to load the climate data and figure out what I want to calculate from that. I have monthly data but I think I need to put it into seasons, so that the plant density is dependent on the previos summer/winter precipitation. That means I have to break it down in to seasons and years. the data collection were all in April so I think that  should end my year in April.
    
```{r}

# Prevent scientific notation
options(scipen = 999)

gbclim <- read.csv("Input files/PRISM_data_Rush.csv")
head(gbclim)

# Convert Date column to Date format
gbclim$Date <- as.Date(paste0(gbclim$Date, "-01"), format = "%Y-%m-%d")

# Extract Year and Month
gbclim$Year <- format(gbclim$Date, "%Y")
gbclim$Month <- format(gbclim$Date, "%m")

# Adjust the Year for water year
gbclim$WaterYear <- as.numeric(gbclim$Year)
gbclim$WaterYear[gbclim$Month %in% c("01", "02", "03")] <- gbclim$WaterYear[gbclim$Month %in% c("01", "02", "03")] - 1

# Calculate total annual precipitation from April to March
total_annual_precip <- aggregate(gbclim$ppt..mm., by = list(Year = gbclim$WaterYear), sum)
mean_annual_temp <- aggregate(gbclim$tmean..degrees.C., by = list(Year = gbclim$WaterYear), mean)

# Calculate seasonal precipitation
gbclim$Season <- cut(as.numeric(gbclim$Month), 
                 breaks = c(0, 2, 4, 8, 10, 12), # Set seasons winter Nov to Feb, spring Mar and Apr, summer May to August, fall sept to oct
                 labels = c("Winter", "Spring", "Summer", "Fall", "Winter"))

seasonal_precip <- aggregate(gbclim$ppt..mm., by = list(Year = gbclim$WaterYear, Season = gbclim$Season), sum)
seasonal_temp <- aggregate(gbclim$tmean..degrees.C., by = list(gbclim$WaterYear, Season = gbclim$Season), mean)

seasonal_temp$x = round(seasonal_temp$x, 2)

colnames(total_annual_precip) <- c("Year", "TAP")
colnames(seasonal_precip) <- c("Year", "Season", "SP")
colnames(mean_annual_temp) <- c("Year", "MAT")
colnames(seasonal_temp) <- c("Year","Season" ,"MST")


# Plot precipitation
ggplot(data = total_annual_precip, aes(x= Year, y= TAP))+
  geom_line()

ggplot(data = seasonal_precip,  aes(x= Year, y= SP, colour = Season))+
  geom_line()

# Plot temperature

ggplot(data = mean_annual_temp, aes(x =Year,  y= MAT))+
  geom_line()

ggplot(data = seasonal_temp, aes(x= Year, y = MST, color = Season))+
  geom_line()

# Now we got to add these to my gbdata. The way I've set it up the total annual ppt for 2013 in gb data needs ppt from April 2012 to March 2013. This should be correct as the year comes in

# I'll start by merging ppt and temp that I created

annual <- merge(total_annual_precip, mean_annual_temp, by = "Year")
seasonal <- merge(seasonal_precip, seasonal_temp, by = c("Year", "Season"))

# Seasonal I'll need to create a column for each season

head(seasonal)
seasonal <- seasonal |> pivot_wider(names_from = Season, values_from = c(SP, MST), names_sep = "_")

```


    Now to merge the climate data with the rest
    
```{r}

colnames(gbdata)[1] <- "Year" # Change name to match clim

gbdata <- merge(gbdata, annual, by = "Year")
gbdata <- merge(gbdata, seasonal, by = "Year")

summary(gbdata$simp_div)
```
Calculate lagged effects of climate and rodents

```{r}

head(annual)
head(seasonal)

lag.annual <- annual
lag.annual$Year <- annual$Year+1

colnames(lag.annual) <- c("Year", "lag.TAP", "lag.MAT")

gbdata <- merge(gbdata, lag.annual, by = "Year")

lag.seasonal <- seasonal

lag.seasonal$Year <- seasonal$Year+1

colnames(lag.seasonal) <- c("Year", "lag.SP_Spring", "lag.SP_Winter",
                            "lag.SP_Fall", "lag.SP_Summer",
                            "lag.MST_Spring", "lag.MST_Winter",
                            "lag.MST_Fall", "lag.MST_Summer")


gbdata <- merge(gbdata, lag.seasonal, by = "Year")

```

Add Cover
```{r}

head(gbdata)

```



Output the final csv

```{r}

write.csv(gbdata, "GreatBasin_Data.csv")

```

