---
title: "ML trials"
author: "R Pienaar"
date: "`r Sys.Date()`"
output: word_document
---

```{r Packages}

library(tidyverse)
library(ggplot2)

```



```{r Load data}

moj.df <-read.csv("Lytle_Veg_Density_13_to_23.csv")
rush.df <- read.csv("Rush_Veg_Density_13_to_23.csv")

```

Exploratory Data Analysis
```{r EDA}

#### Mojave ####

moj.df <- moj.df |> select(DATE, BLOCK, PLOT, BURNED, RODENTS, TRANSECT, SPECIES, MEAN, Density)
head(moj.df)
summary(moj.df$PLOT)
moj.df$MEAN <- as.numeric(moj.df$MEAN)
moj.df$Density <- as.numeric(moj.df$Density)
moj.df$PLOT <- as.factor(moj.df$PLOT)

moj.df[moj.df$PLOT=="UN ",]

ggplot(data = moj.df, aes(PLOT, Density))+
  stat_summary(fun = "sum", geom = "bar")

#### Great Basin ####

rush.df <- rush.df |> select(DATE, Block, Plot,Burned, Rodents, Transect, Plant.Species, mean, Density)
head(rush.df)

```

Cleaning the dataframes
What is my response variable that I am going to measure? Biodiversity indices, Individual Plant species
What are my explanatory variables? Climatic variables, Year, Block, Plot
How does the dataframe need to be formatted?
  Each response needs a unique row of year, block, plot
  We will need unique dataframes for each response
